\chapter{The Platform}


\section{Introduction to rafMetrics platform}
This theoretical work includes a proof-of-concept software stack including various metrics evaluation for classic and modern computer programs. The software stack is further referred as the "Platform".


The metrics that require data acquisition are based on web crawlers interrogations (as described in WebMonitoring tool), experimental data (gathered by computing computer algorithms for various input size in cloud-based systems) or symbolic calculus(for computing rComplexity calculus).

Furthermore, ML-based system is used for estimating rComplexity in the two cases, with known/unknown Big-Theta asymptotic behavior. 

The following metrics are defined inner WebMonitoring Tool:
\begin{itemize}
	\item Resource Crunch (known as Resource, part of WebMonitioring Tool): Average Response Time (daily, weekly, monthly, custom), Average Response Size, Lowest, Medium, Median, Highest metric time, Efficiency Metric	
	\item Website Crunch (known as Website, part of WebMonitioring Tool): Average Response Time (daily, weekly, monthly, custom), Average Response Size (daily, weekly, monthly, custom)
	\item Statistics Resource Manager: Total time, Total number of Requests, Average Time, Average Size, Standard deviation acquired for last 24 hours or all time.
	\item Statistics Website Manager: Total time, Total number of Requests, Average Time, Average Size, Standard deviation acquired for last 24 hours or all time.

	

\end{itemize}


\section{Codebase}
All the code can be accessed and used via GitHub at \href{https://github.com/raresraf/rafMetrics}{https://github.com/raresraf/rafMetrics}.
\\
rafMetrics is licensed under the MIT License: a short and simple permissive license with conditions only requiring preservation of copyright and license notices. Licensed works, modifications, and larger works may be distributed under different terms and without source code.
\\
An easy-to-use, dockerized implementation, can be found at DockerHub within raresraf repositories.


\section{Components}

\subsection{WebMonitoring}
A tool for monitoring multiple network resources and websites. 
Gather data by periodically monitoring specific resources and websites and stores results in database.

\subsubsection{ResourceManager}
Monitors all resources by periodically (timer set default at 1 hour interval) sending requests to existing resources.
Store simple metrics like total time or total requests answer as entries in DB. 

\subsubsection{WebsiteManager}
Monitors all websites by periodically (timer set default at 1 hour interval) generating a HAR (HTTP-Archieve data performance file) for loading metrics corresponding to a website, with Chrome using Browsermob-Proxy.
Also parse and store valuable insights resulted from the HAR file into DB. 
The service uses [speedprofile](https://github.com/parasdahal/speedprofile) engine.

\subsubsection{WebMonitoring API}
Provide an API for interrogating useful metrics from DB.

\subsection{Login}
Backend implementation to provide a simple authentication, registration and management for users inside rafMetrics platform.

\subsection{KubernetesConfig}
Keeps track of all k8s settings

\subsection{DockerConfig}
Keeps track of all Docker Compose/Docker Swarm settings and configurations

\subsection{MySQL}
Database used to store persistent data required by Login and WebMonitoring.
All relations are kept in **Boyce-Codd Normal Form**.

\subsection{deploy\_repo.sh}
Simple script to ensure dockerize and deployment in Kubernetes for all backend components

\subsection{metricsUI}
Frontend implementation of rafMetrics platform based on Flatlogic Template: React Material Admin — Material-UI Dashboard
